---
title: "EricFriesth_hw3"
author: "Eric Friesth"
date: "10/15/2019"
output: html_document
---

## Homework 3
## https://github.com/ericf8/ds202_hw3

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1a. Arrival delay of two or more hours
```{r}

data(flights,package="nycflights13")

LongDelay = subset(flights, flights$arr_delay >= 200)
head(LongDelay)

```
1b. Flew to Houston (IAH or HOU)
```{r}

HoustonFlights = subset(flights, flights$dest == "IAH" | flights$dest == "HOU")
head(HoustonFlights)

```
1c. Departed in summer months (July, August, September)
```{r}

SummerDepart = subset(flights, flights$month == 7 | flights$month == 8 | flights$month == 9)
head(SummerDepart)

```
1d. Arrived more than two hours late, but didn't leave late
```{r}
LongDelay2 = subset(flights, flights$dep_delay <= 0 & flights$arr_delay >= 120)
head(LongDelay2)
```
1e. Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}

QuickFlight = subset(flights, flights$dep_delay >= 60 & flights$dep_delay - flights$arr_delay > 30)
head(QuickFlight)

```
2. This definition of cancelled flights isn't optimal because it will show you a large array of 'True's and 'False's as this operation is just testing for logic. The most import column for finding cancelled flights would be 'dep_delay' or Departure Delay as a plane that has left the airport is considered a "flight."

```{r}

cancelledFlights = is.na(flights$dep_delay) | is.na(flights$arr_delay)
head(cancelledFlights)

```
3. According to the distribution below containing flights that don't have delays grouped by hour, the best time to fly would be between the times of 5:00am and 10:00am.

```{r}

library(ggplot2)

zeroDelay = subset(flights, flights$dep_delay <= 0 & flights$arr_delay <= 0) #created a subset of flights containing only flights with 0 delay or early departures/arrivals

ggplot(zeroDelay, aes(x = zeroDelay$dep_time, group = zeroDelay$hour, bins = 24)) + geom_histogram() + labs(title = "Distribution of Flights with No Delay by Hour", x = "Departure Time (Hour)", y = "Frequency of Flight") + scale_x_continuous(limits=c(0, 2400))


```
4.
flights connects to planes through a single variable, tailnum.
flights connects to airlines through the variable, carrier.
flights connects to airports through the variables origin and dest.
flights connects to weather through the variables year, month, day, hour, origin, 

```{r}

data(airlines,package='nycflights13')
data(airports,package='nycflights13')
data(planes,package='nycflights13')
data(weather,package='nycflights13')

```

```{r}

library(nycflights13)
?airports

```
5.

6.

```{r}

flight_weather = 
flights %>%
inner_join(weather, by = c(
  "origin" = "origin",
  "year" = "year",
  "month" = "month",
  "day" = "day",
  "hour" = "hour"
 ))

head(flight_weather)

weatherDelays = filter(flight_weather, (dep_delay > 60) | (arr_delay) > 60)

head(weatherDelays)


ggplot(weatherDelays, aes(x = weatherDelays$wind_speed)) + geom_histogram(bins = 15) + labs(title = "Distribution of Frequency of Flights by Different Wind Speeds", x = "Wind Speed (Miles per Hour)", y = "Frequency of Flights")

ggplot(weatherDelays, aes(x = weatherDelays$visib)) + geom_histogram(bins = 10) + labs(title = "Distribution of Frequency of Flights by Different Levels of Visibility", x = "Visibility (miles)", y = "Frequency of Flights")

ggplot(weatherDelays, aes(x = weatherDelays$precip)) + geom_histogram(bins = 15) + labs(title = "Distribution of Frequency of Flights by Different Levels of Precipitation", x = "Precipitation (inches)", y = "Frequency of Flights")

```

7. 

```{r}

flight_weather_left = 
flights %>%
left_join(weather, by = c(
  "origin" = "origin",
  "year" = "year",
  "month" = "month",
  "day" = "day",
  "hour" = "hour"
 ))

head(flight_weather_left)

weatherDelays_left = filter(flight_weather_left, (dep_delay > 60) | (arr_delay) > 60)

head(weatherDelays_left)


ggplot(weatherDelays_left, aes(x = weatherDelays_left$wind_speed)) + geom_histogram(bins = 15) + labs(title = "Distribution of Frequency of Flights by Different Wind Speeds", x = "Wind Speed (Miles per Hour)", y = "Frequency of Flights")

ggplot(weatherDelays_left, aes(x = weatherDelays_left$visib)) + geom_histogram(bins = 10) + labs(title = "Distribution of Frequency of Flights by Different Levels of Visibility", x = "Visibility (miles)", y = "Frequency of Flights")

ggplot(weatherDelays_left, aes(x = weatherDelays_left$precip)) + geom_histogram(bins = 15) + labs(title = "Distribution of Frequency of Flights by Different Levels of Precipitation", x = "Precipitation (inches)", y = "Frequency of Flights")

```

8.

```{r}

planes = mutate(planes, "age" = 2019 - planes$year)

flights_planes =
  planes %>% 
  inner_join(flights, by = c(
    "tailnum" = "tailnum"
  ))

head(flights_planes)

ggplot(flights_planes, aes(x = age, y = dep_delay)) + geom_point() + labs()


```






























