---
title: "EricFriesth_hw3"
author: "Eric Friesth"
date: "10/15/2019"
output: html_document
---

## Homework 3
## https://github.com/ericf8/ds202_hw3

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1a. Arrival delay of two or more hours
```{r}

data(flights,package="nycflights13")

LongDelay = subset(flights, flights$arr_delay >= 200)
head(LongDelay)

```
1b. Flew to Houston (IAH or HOU)
```{r}

HoustonFlights = subset(flights, flights$dest == "IAH" | flights$dest == "HOU")
head(HoustonFlights)

```
1c. Departed in summer months (July, August, September)
```{r}

SummerDepart = subset(flights, flights$month == 7 | flights$month == 8 | flights$month == 9)
head(SummerDepart)

```
1d. Arrived more than two hours late, but didn't leave late
```{r}
LongDelay2 = subset(flights, flights$dep_delay <= 0 & flights$arr_delay >= 120)
head(LongDelay2)
```
1e. Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}

QuickFlight = subset(flights, flights$dep_delay >= 60 & flights$dep_delay - flights$arr_delay > 30)
head(QuickFlight)

```
2. This definition of cancelled flights isn't optimal because it will show you a large array of 'True's and 'False's as this operation is just testing for logic. The most import column for finding cancelled flights would be 'dep_delay' or Departure Delay as a plane that has left the airport is considered a "flight."

```{r}

cancelledFlights = is.na(flights$dep_delay) | is.na(flights$arr_delay)
head(cancelledFlights)

```
3. According to the distribution below containing flights that don't have delays grouped by hour, the best time to fly would be between the times of 5:00am and 10:00am.

```{r}

library(ggplot2)

zeroDelay = subset(flights, flights$dep_delay <= 0 & flights$arr_delay <= 0) #created a subset of flights containing only flights with 0 delay or early departures/arrivals

ggplot(zeroDelay, aes(x = zeroDelay$dep_time, group = zeroDelay$hour, bins = 24)) + geom_histogram() + labs(title = "Distribution of Flights with No Delay by Hour", x = "Departure Time (Hour)", y = "Frequency of Flight") + scale_x_continuous(limits=c(0, 2400))


```
4.
flights connects to planes through a single variable, tailnum.
flights connects to airlines through the variable, carrier.
flights connects to airports through the variables origin and dest.
flights connects to weather through the variables year, month, day, hour, origin, 

```{r}

data(airlines,package='nycflights13')
data(airports,package='nycflights13')
data(planes,package='nycflights13')
data(weather,package='nycflights13')

```

```{r}

library(nycflights13)
#?airports

```
5. The airports set is connected through the flights set by the variables origin and dest and thus connects to the weather set through the same variable origin.

6. Based on the three distribution plots below looking at the features, wind speed, precipitation, and visiblity, you are likely to fly in a multitude of wind speed with that distribution being a pretty even bell curve but are unlikely to fly in any kind of low visibility and any kind of precipitation.

```{r}

library(dplyr)
library(tidyverse)

flight_weather = 
flights %>%
inner_join(weather, by = c(
  "origin" = "origin",
  "year" = "year",
  "month" = "month",
  "day" = "day",
  "hour" = "hour"
 ))

head(flight_weather)

weatherDelays = filter(flight_weather, (dep_delay > 60) | (arr_delay) > 60)

head(weatherDelays)


ggplot(weatherDelays, aes(x = weatherDelays$wind_speed)) + geom_histogram(bins = 15) + labs(title = "Distribution of Frequency of Flights by Different Wind Speeds", x = "Wind Speed (Miles per Hour)", y = "Frequency of Flights")

ggplot(weatherDelays, aes(x = weatherDelays$visib)) + geom_histogram(bins = 10) + labs(title = "Distribution of Frequency of Flights by Different Levels of Visibility", x = "Visibility (miles)", y = "Frequency of Flights")

ggplot(weatherDelays, aes(x = weatherDelays$precip)) + geom_histogram(bins = 15) + labs(title = "Distribution of Frequency of Flights by Different Levels of Precipitation", x = "Precipitation (inches)", y = "Frequency of Flights")

```

7. Using the function INNER_JOIN on two datasets returns the rows when there's match ups in those datasets while the function LEFT_JOIN returns all the observations from the left dataset even if they aren't found matched up in the other dataset. 

```{r}


flight_weather_left = 
flights %>%
left_join(weather, by = c(
  "origin" = "origin",
  "year" = "year",
  "month" = "month",
  "day" = "day",
  "hour" = "hour"
 ))

head(flight_weather_left)

weatherDelays_left = filter(flight_weather_left, (dep_delay > 60) | (arr_delay) > 60)

head(weatherDelays_left)


ggplot(weatherDelays_left, aes(x = weatherDelays_left$wind_speed)) + geom_histogram(bins = 15) + labs(title = "Distribution of Frequency of Flights by Different Wind Speeds", x = "Wind Speed (Miles per Hour)", y = "Frequency of Flights")

ggplot(weatherDelays_left, aes(x = weatherDelays_left$visib)) + geom_histogram(bins = 10) + labs(title = "Distribution of Frequency of Flights by Different Levels of Visibility", x = "Visibility (miles)", y = "Frequency of Flights")

ggplot(weatherDelays_left, aes(x = weatherDelays_left$precip)) + geom_histogram(bins = 15) + labs(title = "Distribution of Frequency of Flights by Different Levels of Precipitation", x = "Precipitation (inches)", y = "Frequency of Flights")

```

8. Using the scatter plot below, I see that there isn't an overwhelming correlation between the age of a plane and departure delays but we do see more incidents of total flights in planes younger than 40.

```{r}

planes = mutate(planes, "age" = 2019 - planes$year)

flights_planes =
  planes %>% 
  inner_join(flights, by = c(
    "tailnum" = "tailnum"
  ))

head(flights_planes)

ggplot(flights_planes, aes(x = age, y = dep_delay)) + geom_point() + labs()


```

9. In the first argument, we are shown the flights dataset with the faa from airports connected by dest 
and in the second, we are shown the airports dataset with the dest of flights connected by faa. 

```{r}

anti_join(flights, airports, by = c("dest" = "faa"))

anti_join(airports, flights, by = c("faa" = "dest"))

```





























